Resonant Scattering in a Uniform Sphere with Large Optical Depth

### 2021-10-05 ms.tex
corrective term -> correction in abstract

The solution of the radiative transfer equation for resonant scattering of Lyman $\alpha$ photons is explored in spherical geometry. A monochromatic source of photons is assumed at the center of a uniform sphere of constant gas density. The intensity within the sphere and at its surface is found in the limit of large line-center optical depth using the Eddington approximation for the angular dependence. Two specific limits are studied: a steady source and an impulsive source. For the steady-state problem, the solution can be represented as a sum of two terms: a previously-known analytic solution of the equation with the mean intensity $J=0$ at the surface, and a novel, semi-analytic solution which enforces the correct zero-ingoing intensity at the surface. This solution is compared to that of the Monte-Carlo method, which is valid at arbitrary optical depth. It is shown that the size of corrective term depends on the line center optical depth, and may explain discrepancies in previous studies that used a $J=0$ boundary condition. For the impulsive problem, the time, spatial, and frequency dependence of the diffusion equation are expressed using an eigenfunction expansion and solved numerically to characterize the escape time distribution and spectra of photons initialized at the center of the sphere. It is shown that the lowest-order eigenfrequency of the semi-analytic calculation agrees well with the decay constant of the Monte Carlo escape time distribution at line-center optical depths above ${\sim} 10^7$. A fitting function for the distribution is found which can be sampled to accelerate Monte Carlo radiation transfer calculations at large optical depths.

>>>>

The solution of the radiative transfer equation for resonant scattering of Lyman $\alpha$ photons is explored in spherical geometry. A monochromatic source of photons is assumed at the center of a uniform sphere of constant gas density. The intensity within the sphere and at its surface is found in the limit of large line-center optical depth using the Eddington approximation for the angular dependence. Two specific limits are studied: a steady source and an impulsive source. For the steady-state problem, the solution can be represented as a sum of two terms: a previously-known analytic solution of the equation with the mean intensity $J=0$ at the surface, and a novel, semi-analytic solution which enforces the correct zero-ingoing intensity at the surface. This solution is compared to that of the Monte-Carlo method, which is valid at arbitrary optical depth. It is shown that the size of the correction depends on the line center optical depth, and may explain discrepancies in previous studies that used a $J=0$ boundary condition. For the impulsive problem, the time, spatial, and frequency dependence of the diffusion equation are expressed using an eigenfunction expansion and solved numerically to characterize the escape time distribution and spectra of photons initialized at the center of the sphere. It is shown that the lowest-order eigenfrequency of the semi-analytic calculation agrees well with the decay constant of the Monte Carlo escape time distribution at line-center optical depths above ${\sim} 10^7$. A fitting function for the distribution is found which can be sampled to accelerate Monte Carlo radiation transfer calculations at large optical depths.
